---
title: "BIOS 512 Homework 2"
output: pdf_document
---

## Load Necessary Packages...

If you need to install tidyverse, you can use install.packages("tidyverse"). If you already have tidyverse installed, you can just load the package:

```{r, results="hide", message=FALSE, warning=FALSE}
library(tidyverse);
```

## Question 1

What are three ways that we can assign the value of 25 to y?

Answer:

```{r}
y1 = 25
y2 <- 25
assign("y3", 25)
print(y1)
print(y2)
print(y3)
```
## Question 2

How would I print the following line, including the quotes, in R?\
**He said, "I'm Garth Marenghi. Author. Dreamweaver. Visionary. Plus actor."**

Answer:
```{r}
cat('He said, "I\'m Garth Marenghi. Author. Dreamweaver. Visionary. Plus actor."')
```

## Question 3

What if I wanted to add backslashes into the statement from Question 2 to make the statement below?\
**He said, "I'm Garth Marenghi. Author\\Dreamweaver\\Visionary. Plus actor."**

Answer:
```{r}
cat('He said, "I\'m Garth Marenghi. Author\\\\Dreamweaver\\\\Visionary. Plus actor."')
```

## Question 4

Show two ways to get the following array: 1 2 3

Answer:
```{r}
x <- c(1,2,3)
print(x)

y <- 1:3
print (y)
```

## Question 5

What does R call things like +, -, sin(), c(), etc? What about \<-?

Answer:
+, -, sin() and c() are all called funcations. 
<- is called an assignment. 

## Question 6

What's the difference in the way R processes the while() and the for() below?

```{r}
x <- 0;
while (x <= 3) {
    x <- x + 1;
}
x

for(y in c(1,2,3)) {
     y <- y + 1;
}
y
```
Answer: 
The while() loop block is evaluated until x's value is less than or equal to 3. The for() loop evaluates the block many times, with y bound to subsequent values of 
1,2, and 3.

## Question 7

Create the Pythagorean formula and evaluate it with a=3 and b=4. Print the output.

Answer:
```{r}
a <- 3
b <- 4
c <- sqrt(a^2 + b^2)
print(c)
```
```{r}
pythagorean <- function(a, b) {
  sqrt(a^2 + b^2)
}
answer <- pythagorean(3, 4)
print(answer)
```

## Question 8

Load the help for the built in sin() function.

Answer:
```{r}
?sin
```

## Question 9

Which version of the counter function works? What is the difference in the way R processes the two functions?

```{r}
counter1 <- function(start, step){
    val <- start;
    function(){
        old_val <- val;
        val <- val + step;
        old_val;
    }
}
counter_from_1 <- counter1(1,1);
counter_from_1()
counter_from_1()

counter2 <- function(start, step){
    val <- start;
    function(){
        old_val <- val;
        val <<- val + step;
        old_val;
    }
}
counter_from_1 <- counter2(1,1);
counter_from_1()
counter_from_1()
```
Answer:
The second version on the counter works.
In the counter1 function, val is being created in the temporary enviornemnt that appears when counter1 is called. When it assigns val <- val + step, it creates a new local val inside the inner function, but doesn't change the previous val. So, every time you call the function, it starts from the same val and never counts higher.
The counter2 function uses the double arrow which tells R to update val in the environment above so it its updated each time it is called.

## Question 10

a)  Use read_csv the cars.csv and store it in a data frame.
b)  Then, group the data frame by Make, get averages across the numeric variables, and then sort by Volume in descending order. Hint: Use summarise(across(c(<variable list>), mean)) to get the averages.

Answer:
```{r}
library(readr) 
library(dplyr)

df <- read_csv("cars.csv")

numeric_vars <- df %>%
  select(where(is.numeric)) %>%
  colnames()

summary_df <- cars_df %>%
  group_by(Make) %>%
  summarise(across(all_of(numeric_vars), mean, na.rm = TRUE)) %>%
  arrange(desc(Volume))
```

## Question 11

Make a function that returns the Fibonacci sequence, then call it 7 times to return the first 7 values of the sequence. Use the correct counter function from question 9 for inspiration.
